---
title: "Reproducible Research: Peer Assessment 1"
author: "S.Sriranjani"
date: "20 August 2018"
output: html_document
---
&nbsp;

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

###1. Code for reading in the dataset

```{r read}
dt<-read.csv("D:/Coursera/Data Science/activity.csv")
```

###2. Histogram of the total number of steps taken each day

```{r histogram}
d1<-dt %>% group_by(date) %>% summarise(steps=sum(steps))
hist(d1$steps)
```

###3. Mean and median number of steps taken each day

```{r mean}
mean(d1$steps,na.rm=TRUE)
```

```{r median}
median(d1$steps,na.rm=TRUE)
```

###4. Time series plot of the average number of steps taken

```{r time series plot}
d2<-dt %>% group_by(interval) %>% summarise(steps=mean(steps,na.rm=TRUE))
plot(d2$interval,d2$steps,type="l")
```

###5. The 5-minute interval that, on average, contains the maximum number of steps

```{r maximum_steps}
d2[d2$steps==max(d2$steps),]
```

Maximum number of steps is at interval `r d2[d2$steps==max(d2$steps),1]`

###6.Code to describe and show a strategy for imputing missing data
The total number of missing values in the dataset

```{r total_na}
sum(is.na(dt$steps))
```

We would replace the missing values(NAs) with the average of steps taken in the 5-minute interval averaged across all days.

Creating a new data set with the missing data filled in - 

```{r new_dataset}
newdt=merge(dt,d2,by="interval")
newdt[is.na(newdt$steps.x),"steps.x"]<-newdt[is.na(newdt$steps.x),"steps.y"]
newdt$steps<-newdt$steps.x
newdt$steps.x<-NULL
newdt$steps.y<-NULL
```

###7.Histogram of the total number of steps taken each day after missing values are imputed

```{r new_histogram}
newd1<-newdt %>% group_by(date) %>% summarise(steps=sum(steps))
hist(newd1$steps)
```

Mean and median number of steps taken each day after missing values are imputed

```{r new_mean}
mean(newd1$steps)
```

```{r new_median}
median(newd1$steps,na.rm=TRUE)
```
###8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

Adding weekday column to the data, Creating two new datasets - dtwe for weekend data and dtwd for weekdays data

```{r weekday}
newdt$weekday<-weekdays(as.Date(newdt$date))
dtwe<-newdt[newdt$weekday %in% c("Saturday","Sunday"),]
dtwd<-newdt[!(newdt$weekday %in% c("Saturday","Sunday")),]
head(dtwd)
head(dtwe)
```
Panel plot
```{r panel plot}
dtwe<-dtwe %>% group_by(interval) %>% summarise(steps=mean(steps,na.rm=TRUE))
dtwd<-dtwd %>% group_by(interval) %>% summarise(steps=mean(steps,na.rm=TRUE))
par(mfrow=c(1,2))
plot(dtwd$interval,dtwd$steps,type="l",ylim=c(0,250),xlab="Inteval",ylab="Steps")
title("Weekdays")
plot(dtwe$interval,dtwe$steps,type="l",ylim=c(0,250),xlab="Inteval",ylab="Steps")
title("Weekends")
```
